Author: "Tom Bair"
date: ""
output: html_document
---

Inital import of libraries: This is a typical DESeq2 analysis

```{r,eval=TRUE,echo=TRUE,results='hide',message=FALSE}
library("DESeq2")
library("gplots")
library("genefilter")
library("RColorBrewer")
library("gtools")
source("functions/dist.pear.R")
library("knitr")
library("gProfileR")
library("limma")
```


Import the data file and the phenotype file to assign samples to categorical variables. 

```{r,eval=TRUE,cache=FALSE,echo=TRUE,results='hide',message=FALSE}
organism = 'mmusculus'
pheno_file <- ""
counts <- read.table("",header=TRUE,row.names=1,sep='\t')
pheno <- read.table(pheno_file,header=TRUE, sep=",",row.names=1)
project_name <- ""
file_path = paste("./",project_name,"_results/",sep="")
if (!file.exists(file_path)){
  dir.create(file_path, mode = "0777")
}

dds <- DESeqDataSetFromMatrix(countData = counts, colData=pheno, design = ~Type ) #<---insert design column name here
rld <- rlogTransformation(dds)
dds<- DESeq(dds)
print(dim(rld))
distsRL <- dist(t(assay(rld)))
mat <- as.matrix(distsRL)
```

Values assigned to samples ***verify that this is correct ***

```{r, echo=FALSE}
kable(pheno)
```

PCA plot, should show similarites between related samples, however it may be more influenced by other technical factors if the biological differences are slight

```{r, fig.width=8, fig.height=6 }
plotPCA(rld,intgroup = c("Type"))
```

Correlation heatmap, similar concept to the PCA plot just a different technique.

```{r, fig.width=8, fig.height=6}
heatmap.2(mat,trace="none", margin = c(10,10),col = greenred(75),labRow = pheno$Type,labCol = pheno$ID)
```

Compare Template

```{r}
res <- results(dds, pAdjustMethod = "fdr",contrast = c("Type","",""))
resSig <- !is.na(res$padj) & res$padj <= 0.01 & abs(res$log2FoldChange) >= 2
```

Number of significant genes

```{r}
summary(resSig)
```

Heatmap of significant results compairing # add comparison

```{r, fig.width=8, fig.height=6}
heatmap.2(assay(rld)[resSig,],distfun=dist.pear.minus,hclustfun=hclust.ave, col = greenred(75), labCol=pheno$Type,scale="row",dendrogram = "both", trace="none",labRow=FALSE, margin=c(10,6))
  
```

Volcano Plot
```{r, fig.width=8, fig.height=6}
  plot(res$log2FoldChange, -log10(res$pvalue ), ylab="-log10(p-value)", xlab="logFC")
  points(res[resSig,"log2FoldChange"], -log10(res[resSig,"pvalue"]), col=2, pch=20)
```


Write out the data 
```{r}

normcounts <- as.data.frame(counts(dds, normalized=TRUE))
colnames(normcounts)<-paste(colnames(normcounts),pheno$Type)
colnames(res) <- paste(colnames(res),"NIK_DN1_v_CTL",sep="_")
colnames(res2) <- paste(colnames(res),"NIK_HSC_v_CTL",sep="_")
DN1 <- resSig
HSC <- resSig2
d <- cbind(normcounts,DN1,HSC,res,res2)

write.csv(d, quote=TRUE, file='results.csv')
i <- resSig & resSig2
g<- gprofiler(rownames(d[i,]),organism = organism)
write.csv(g,file="go_enrichment.csv")
```

```{r}
  v <- cbind(resSig,resSig2)
  colnames(v)<- c("DN1","HSC")
```

Overlap of significant results

```{r, fig.width=9, fig.height=6}
  vennDiagram(v)
```

Plot out the genes that have significance in both tests

```{r, fig.width=8, fig.height=6}
i <- resSig & resSig2
heatmap.2(assay(rld)[i,],distfun=dist.pear.minus,hclustfun=hclust.ave, col = greenred(75), labCol=pheno$Type,scale="row",dendrogram = "both", trace="none",labRow=FALSE, margin=c(10,6))
  
```


Session info:

```{r}
sessionInfo()
```
