Author: "Tom Bair"
date: "July 19, 2016"
output: html_document
---

Inital import of libraries: This is a typical DESeq2 analysis

```{r,eval=TRUE,echo=TRUE,results='hide',message=FALSE}
library("DESeq2")
library("gplots")
library("genefilter")
library("RColorBrewer")
library("gtools")
source("~/R_functions/dist.pear.R")
library("knitr")
library("gProfileR")
library("limma")
library("AnnotationDbi")
library("org.Mm.eg.db")
library("calibrate")
#columns(org.Mm.eg.db)
```


Import the data file and the phenotype file to assign samples to categorical variables. 

```{r,eval=TRUE,cache=FALSE,echo=TRUE,results='hide',message=FALSE}
organism = 'mmusculus'
pheno_file <- "phenodata.csv"
project_name <- "inital"
counts <- read.table("combined.counts",header=TRUE,row.names=1,sep='\t')
pheno <- read.table(pheno_file,header=TRUE, sep=",",row.names=1)
project_name <- ""
file_path = paste("./",project_name,"_results/",sep="")
if (!file.exists(file_path)){
  dir.create(file_path, mode = "0777")
}

pheno$Type <- paste(pheno$Cell,pheno$Infection,sep="_" )
dds <- DESeqDataSetFromMatrix(countData = counts, colData=pheno, design = ~Type ) #<---insert design column name here
rld <- rlogTransformation(dds)
dds<- DESeq(dds)
print(dim(rld))
distsRL <- dist(t(assay(rld)))
mat <- as.matrix(distsRL)
```

Values assigned to samples ***verify that this is correct ***

```{r, echo=FALSE}
kable(pheno)
```

PCA plot, should show similarites between related samples, however it may be more influenced by other technical factors if the biological differences are slight

```{r, fig.width=8, fig.height=6 }
plotPCA(rld,intgroup = c("Type"))
```

Correlation heatmap, similar concept to the PCA plot just a different technique.

```{r, fig.width=8, fig.height=6}
heatmap.2(mat,trace="none", margin = c(10,10),col = greenred(75),labRow = pheno$Type,labCol = pheno$ID)
```

Compare 4T1_shCtl Naive vs Infected

```{r}
res <- results(dds, pAdjustMethod = "fdr",contrast = c("Type","4T1_shCTL_Naive","4T1_shCTL_Infected"))
res$symbol <- mapIds(org.Mm.eg.db, keys=row.names(res), column="SYMBOL",keytype="ENSEMBL", multiVals="first")
resSig <- !is.na(res$padj) & res$padj <= 0.05
```

Number of significant genes (FDR adjusted p-value < 0.05)

```{r}
summary(resSig)
```

Heatmap of significant results compairing 4T1_shCtl Naive vs Infected # add comparison

```{r, fig.width=8, fig.height=6}
heatmap.2(assay(rld)[resSig,],distfun=dist.pear.minus,hclustfun=hclust.ave, col = greenred(75), labCol=pheno$Type,scale="row",dendrogram = "both", trace="none",labRow=res[resSig,'symbol'], margin=c(10,6))
  
```

Volcano Plot
```{r, fig.width=8, fig.height=6}
  plot(res$log2FoldChange, -log10(res$pvalue ), ylab="-log10(p-value)", xlab="logFC")
  points(res[resSig,"log2FoldChange"], -log10(res[resSig,"pvalue"]), col=2, pch=20)
```

Compare 4T1_shBST_2 Naive vs Infected

```{r}
res2 <- results(dds, pAdjustMethod = "fdr",contrast = c("Type","4T1_shBST_2_Naive","4T1_shBST_2_Infected"))
res2$symbol <- mapIds(org.Mm.eg.db, keys=row.names(res2), column="SYMBOL",keytype="ENSEMBL", multiVals="first")
res2Sig <- !is.na(res2$padj) & res2$padj <= 0.05
```

Number of significant genes

```{r}
summary(res2Sig)
```

Heatmap of significant results compairing # add comparison

```{r, fig.width=8, fig.height=6}
heatmap.2(assay(rld)[res2Sig,],distfun=dist.pear.minus,hclustfun=hclust.ave, col = greenred(75), labCol=pheno$Type,scale="row",dendrogram = "both", trace="none", labRow=res2[res2Sig,'symbol'],margin=c(10,6))
  
```

Volcano Plot
```{r, fig.width=8, fig.height=6}
  plot(res2$log2FoldChange, -log10(res2$pvalue ), ylab="-log10(p-value)", xlab="logFC")
  points(res2[res2Sig,"log2FoldChange"], -log10(res2[res2Sig,"pvalue"]), col=2, pch=20)
  textxy(res2[res2Sig,"log2FoldChange"], -log10(res2[res2Sig,"pvalue"]), labs=res2[res2Sig,"symbol"])
```


Write out the data 
```{r}

normcounts <- as.data.frame(counts(dds, normalized=TRUE))
colnames(normcounts)<-paste(colnames(normcounts),pheno$Type)
write.table(res[resSig,],file="shCrl_N_v_I.txt",sep="\t",quote=FALSE)
write.table(res2[res2Sig,],file="shBST_N_v_I.txt",sep="\t",quote=FALSE)
colnames(res) <- paste(colnames(res),"shCtl_N_v_I",sep="_")
colnames(res2) <- paste(colnames(res2),"shBST_N_v_I",sep="_")

 shCtl<- resSig
shBST <- res2Sig
d <- cbind(normcounts,shCtl,shBST,res,res2)

write.csv(d, quote=TRUE, file='results.csv')
i <- resSig | res2Sig
g<- gprofiler(d[i,"symbol_shCtl_N_v_I"],organism = organism)
write.csv(g,file="go_enrichment.csv")
```

```{r}
  v <- cbind(resSig,res2Sig)
  colnames(v)<- c("shCtl","shBST")
```

Overlap of significant results

```{r, fig.width=9, fig.height=6}
  vennDiagram(v)
```

Plot out the genes that have significance in either tests

```{r, fig.width=8, fig.height=6}
i <- resSig | res2Sig
heatmap.2(assay(rld)[i,],distfun=dist.pear.minus,hclustfun=hclust.ave, col = greenred(75), labCol=pheno$Type,scale="row",dendrogram = "both", trace="none",labRow=FALSE, margin=c(10,6))
  
```


Session info:

```{r}
sessionInfo()
```
